Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-02-07 draft.tex
move dataset section and rename sections
expand introduction and related works

	tension \cite{veerasamy_rising_2022,albahar_deepfakes_2005}. The field of Explainable AI (XAI)
	shows promise for producing useful and interpretable results from such models. Explainable AI
	refers to the ability of artificial intelligence systems, such as machine learning models and
	neural networks, to provide understandable and interpretable explanations for their decisions
	or predictions \cite{hind_explaining_2019}. This means that XAI systems can articulate why
	they made a particular decision, what factors influenced it and how confident they are in
	their conclusion \cite{hind_explaining_2019}. For the detection of audio deepfakes, this could
	mean that the model can justify its classification with feature-based evidence, allowing for
	verification of the result by the end user. Much of the existing research, further discussed
	in the next section, follows the path of producing a model with the best possible benchmarks
	against datasets of samples, without taking into account if the results can be usefully
	interpreted, while existing explorations into making audio deepfake detection explainable have
	produced results that often require a lot of background knowledge to meaningfully interpret.
	Aside from this, explainability in the area of audio deepfake detection remains an open
	challenge \cite{cuccovillo_open_2022}, one even promoted by regulatory bodies such as the
	European Union with the "right to explain" \cite{goodman_european_2017}. With the context of
	the research discussed in the next section, we would like to use new combinations of features,
	most of which are perceptible, with a novel model architecture to generate explanations for
	fake or real audio classification, with the goal of singling out features that are useful not
	only for an accurate classification but also for an explanation that can be understood by the
	average person, explained using features that can be heard by the human ear.

@@

	This section will summarize some efforts in previous research categorized by the use of
	perceptible features or attempts to create explainable models.
	\subsection{Perceptible Features}

@@

	\section{Experiments}

@@

	\section{Results}

>>>>

	tension \cite{veerasamy_rising_2022,albahar_deepfakes_2005}.
	The field of Explainable AI (XAI) shows promise for producing useful and interpretable results
	from such models. Explainable AI refers to the ability of artificial intelligence systems,
	such as machine learning models and neural networks, to provide understandable and
	interpretable explanations for their decisions or predictions \cite{hind_explaining_2019}.
	This means that XAI systems can articulate why they made a particular decision, what factors
	influenced it and how confident they are in their conclusion \cite{hind_explaining_2019}. For
	the detection of audio deepfakes, this could mean that the model can justify its
	classification with feature-based evidence, allowing for verification of the result by the end
	user. Much of the existing research, further discussed in the next section, follows the path
	of producing a model with the best possible benchmarks against datasets of samples, without
	taking into account if the results can be usefully interpreted, while existing explorations
	into making audio deepfake detection explainable have produced results that often require a
	lot of background knowledge to meaningfully interpret. Aside from this, explainability in the
	area of audio deepfake detection remains an open challenge \cite{cuccovillo_open_2022}, one
	even promoted by regulatory bodies such as the European Union with the "right to explain"
	\cite{goodman_european_2017}.
	In this work, we will make a distinction of two main categories of input features for a
	synthetic voice detection model: perceptible and imperceptible features. Perceptible features
	are features that can be perceived by the human ear, often vocal qualities such as jitter,
	shimmer or pitch fluctuation that have a wide range of uses even outside of audio
	classification such as diagnosis of disease \cite{chaiwongyen_deepfake-speech_2023}, while
	imperceptible features are typically out of the range of human hearing, and may not be
	directly reflect a vocal quality, otherwise referred to as "speaker-independent" features
	\cite{liu_hidden--wave_2023}. We use a new combination of features for our method, with an
	emphasis being placed on perceptible features, as we believe these to be the most likely to be
	understood by the layperson.
	We then combine these features together using a novel model architecture, with separate
	sub-models for each input feature that are combined and processed in a final model, referred
	to as the terminus model, allowing us to process different kinds of features separately and
	avoid unnecessary, costly preprocessing.
	Explanations are generated using the LIME method \cite{ribeiro_why_2016}, which allows us to,
	for an individual classification assess the impact of each input feature on the final result.
	In order to assess how useful the average local explanation is to a potential end user, and
	due to a lack of such assessments in previous research, we also introduce two metrics for
	aggregate evaluation of large number of generated LIME explanations, which we use to make a
	conclusion about the value of the selected feature set for the purpose of sensible
	explanations.

@@

	This section will summarize some efforts in previous research divided into categories of
	perceptible and imperceptible features, as well as previous efforts at explainability in this
	field.
	\subsection{Synthetic Voice Detection With Imperceptible Features}
	The most common approach in this field makes use of either learned or hand-crafted
	imperceptible features. Examples include spectrographic features such as the mel-spectrogram
	and its hand-crafted derivative the mel-frequency cepstral coefficients (MFCCs), both of which
	are present in our method due to their widespread successful use.
	The work of Anagha et al. \cite{anagha_audio_2023} is an example that makes use of
	mel-spectrograms in combination with convolutional neural networks, achieving well-performing
	results on the ASVSpoof2019 dataset \cite{wang_asvspoof_2020}.
	A further work \cite{yan_initial_2022} makes use of MFCCs, in combination with other
	hand-crafted imperceptible features such as linear frequency cepstral coefficients (LFCCs) to
	not only classify audio samples as synthetic or authentic, but also to correctly classify the
	vocoder used to produce the sample with nearly perfect accuracy on a handcrafted dataset.
	Other researchers, such as in the work of Qais et al. \cite{qais_deepfake_2022} make use of
	Fourier transforms, such as the short term Fourier transform (STFT) to achieve accurate
	reuslts on the ASVSpoof2017 dataset \cite{delgado_asvspoof_2018}.
	\subsection{Synthetic Voice Detection With Perceptible Features}
	There is also a smaller but still notable body of work encompassing the use of perceptible
	features for the detection of synthetic voices, making use of a variety of features, some of
	which were selected for use in our experiments.

@@

	\section{Method}

@@

	\section{Experimental Results}
